---
title: "Final Project"
author: "Pizon Shetu"
date: "11:59PM May 24, 2020"
output:
  word_document: default
  pdf_document: default
---

Libraries used in my project

```{r}
install.packages('stringr', repos = c('http://rforge.net', 'http://cran.rstudio.org'),
                 type = 'source')
getOption("repos")
library(ggplot2)
pacman::p_load(dplyr, tidyr, magrittr, mlr, missForest)
library(ggmap)
library(Hmisc)
library(psych)
library(car)
library(readr)
library(varhandle)
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
library(ipred)       # bagging
library(caret)
library(resample)
# library(YARF)
# pacman::p_install_gh("kapelner/YARF", subdir = "YARF", ref = "dev")
# pacman::p_install_gh("kapelner/YARF", subdir = "YARFJARs", ref = "dev")
# pacman::p_install_gh("kapelner/YARF", subdir = "YARF", ref = "dev")
# pacman::p_load(YARF)
# pacman::p_load(YARF)
# 
# install.packages("rJava", type = "source")
# install.packages("rmarkdown")
# getOption("repos")
# update.packages()
# 
# options(java.parameters = "-Xmx4000m")
# pacman::p_load(rJava)
# 
# .jinit()
# 
# 
# 
# 
# pacman::p_install_gh("kapelner/YARF/YARF", ref = "dev")
housing.data = read.csv("housing_data_2016_2017.csv")
#housing.data$sale_price
#head(housing.data)
```

View the data and the structure of the data to see what we're working with

```{r}
glimpse(housing.data)
#2,230 obserations and 55 variables
#summary(housing.data)
na.vec = which(!complete.cases(housing.data$sale_price))

housing.data = housing.data[-na.vec,]
sapply(housing.data[,1:55], function(x) sum(is.na(x)))


```
There are alot of variables in this dataset, so I will be dropping the one's which do not relate to the price using dplyr

```{r}
house = select(housing.data, -c(URL,WorkerId,AssignmentId,HITId,HITTypeId,Title,Description,Keywords,Reward,CreationTime,MaxAssignments,RequesterAnnotation,AssignmentDurationInSeconds,AutoApprovalDelayInSeconds,url,Expiration,NumberOfSimilarHITs,LifetimeInSeconds,AssignmentId,WorkerId,AssignmentStatus,AcceptTime,SubmitTime,AutoApprovalTime,ApprovalTime,RejectionTime,RequesterFeedback,WorkTimeInSeconds,LifetimeApprovalRate,Last30DaysApprovalRate,Last7DaysApprovalRate,date_of_sale,listing_price_to_nearest_1000,pct_tax_deductibl,model_type,num_half_bathrooms,full_address_or_zip_code ))



glimpse(house)
#We have dropped 25 variables but our observations are still the same
#When cleaning and dealing with missing data i realized many of the features just had too much missing data thus for better prediction it was more efficient to drop those fetures
```






```{r}
#Need to re-structure some of the data-types which are factors to numerical and fixe some missingness


house$dogs_allowed = ifelse(house$dogs_allowed == "no", 0, 1)

house$cats_allowed = ifelse(house$cats_allowed == "no", 0, 1)

house$dining_room_type = factor(house$dining_room_type, ordered = FALSE)

house$fuel_type = factor(house$fuel_type, ordered = FALSE)

house$kitchen_type = factor(house$kitchen_type, ordered = FALSE)

house$coop_condo = factor(tolower(house$coop_condo))

house$garage_exists = ifelse(is.na(house$garage_exists), 0, 1)


house$num_floors_in_building = as.integer(house$num_floors_in_building)
house$num_bedrooms = as.integer(house$num_bedrooms)
house$num_full_bathrooms = as.integer(house$num_full_bathrooms)
house$num_total_rooms = as.integer(house$num_total_rooms)

house$sale_price = as.character(house$sale_price)
house$sale_price = gsub("[^0-9.-]", "", house$sale_price)
house$sale_price = as.numeric(house$sale_price)

house$maintenance_cost = as.character(house$maintenance_cost)
house$maintenance_cost = gsub("[^0-9.-]", "", house$maintenance_cost)
house$maintenance_cost = as.numeric(house$maintenance_cost)

house$common_charges = as.character(house$common_charges)
house$common_charges = gsub("[^0-9.-]", "", house$common_charges)
house$common_charges = as.numeric(house$common_charges)

house$parking_charges = as.character(house$parking_charges)
house$parking_charges = gsub("[^0-9.-]", "", house$parking_charges)
house$parking_charges = as.numeric(house$parking_charges)

house$total_taxes = as.character(house$total_taxes)
house$total_taxes = gsub("[^0-9.-]", "", house$total_taxes)
house$total_taxes = as.numeric(house$total_taxes)

house$dining_room_type[house$dining_room_type == "none" & !is.na(house$dining_room_type)] = "other"
house$dining_room_type[house$dining_room_type == "dining area" & !is.na(house$dining_room_type)] = "other"
house$fuel_type[house$fuel_type == "none" & !is.na(house$fuel_type)] = "other"
```

Now lets take care of missing values
```{r}
# Shows the number of missing values for each columns 
colSums(sapply(house, is.na))


#Isolating the zipcode from the address which takes away too much variability in the choices while keep a zip code which can be associated with a certain area
#Turns out full address did not help with our errors
# rightSubStr = function(x,n){
#   substr(x, nchar(x)-n+1, nchar(x))
# }
# house$full_address_or_zip_code = gsub("[^0-9.-]", "", house$full_address_or_zip_code)
# house$full_address_or_zip_code = as.character(house$full_address_or_zip_code)
# house$full_address_or_zip_code = rightSubStr(house$full_address_or_zip_code, 5)
# house$full_address_or_zip_code = as.integer(house$full_address_or_zip_code)
# house$full_address_or_zip_code
# class(house$full_address_or_zip_code)


pacman::p_load(nycflights13, tidyverse, magrittr, data.table)



missing = tbl_df(apply(is.na(house), 2, as.numeric))
colnames(missing) = paste("is_missing_", colnames(house), sep = "")
missing %<>%
  select_if(function(x){sum(x) > 0})
missing
house_imp = missForest(data.frame(house))$ximp
house_imp

sd(house_imp$sq_footage)
final_house_df = cbind(house_imp,missing)
#Show the number of missing value for all variable in combined dataset 

#final_house_df
#sapply(final_house_df[,1:19], function(x) sum(is.na(x)))

#glimpse(final_house_df)

#Time to split into training and test 

set.seed(123)
split <- sample(seq_len(nrow(house_imp)), size = floor(0.7 * nrow(house_imp)))
train <- house_imp[split, ]
test <- house_imp[-split, ]
dim(train)
dim(test)

```

Lets see the distribution in housing prices
```{r}
#The type for price is factor so we make it to numeric as I find it easier to work with as well as plotting with numeric is easier
price = house_imp$sale_price

price = price / 1000


hist(price,
     data = house_imp,
     main = 'Price Distribution',
     xlab = 'Price in 100K or times 1000',
     ylab = 'Frequency',
     col = 'green',
     freq = TRUE,
     )

#Now lets see the distribution in bedrooms 
bedrooms = house_imp$num_bedrooms
hist(bedrooms,
     main = 'Bedroom Distribution',
     xlab = 'Bedrooms',
     ylab = 'Frequency',
     col = 'brown',
     
     )


plot(y = house_imp$sale_price/100000, x = house_imp$sq_footage, xlab = 'Sq Footage', ylab = 'Price in 100K', main = 'Price Distribution by Sq Footage', col = 'darkred')

```



Lets try to fit it using regression tree, going to use the package rpart
```{r}



# install.packages('tree')
# library('rpart.plot')
# library(rpart)

tree_model = rpart(sale_price ~ ., data = train)

tree_model
#Not sure how rpart is applying its cost function but apparently rpart performs 10-fold cross validation so that the error associated with a given α value is computed on the hold-out validation data. 
rpart.plot(tree_model)

plotcp(tree_model)

tree_model$cptable

pred <- predict(tree_model, newdata = test)
RMSE(pred = pred, obs = test$sale_price)

pacman::p_load(rpart)


summary(tree_model)
#We will try further tuning by changing minsplit and maxdepth which controls the minimum number of data points and maximum number of internal nodes 

tree_model_2 <- rpart(
    formula = sale_price ~ .,
    data    = train,
    method  = "anova", 
    control = list(minsplit = 10, maxdepth = 30, xval = 10)
)

plotcp(tree_model_2)
abline(v = 9, lty = "dashed")
tree_model_2$cptable
predictionrt <- predict(tree_model_2, newdata = test)
RMSE(pred = pred, obs = test$sale_price)

#We need find the optimal number for our hyper-parameter min-split and max-depth

hyper_grid <- expand.grid(
  minsplit = seq(5, 20, 1),
  maxdepth = seq(8, 15, 1)
)

head(hyper_grid)


nrow(hyper_grid)

models <- list()

for (i in 1:nrow(hyper_grid)) {
  
  # get minsplit, maxdepth values at row i
  minsplit <- hyper_grid$minsplit[i]
  maxdepth <- hyper_grid$maxdepth[i]

  # train a model and store in the list
  models[[i]] <- rpart(
    formula = sale_price ~ .,
    data    = train,
    method  = "anova",
    control = list(minsplit = minsplit, maxdepth = maxdepth)
    )
}


# function to get optimal cp
get_cp <- function(x) {
  min    <- which.min(x$cptable[, "xerror"])
  cp <- x$cptable[min, "CP"] 
}

# function to get minimum error
get_min_error <- function(x) {
  min    <- which.min(x$cptable[, "xerror"])
  xerror <- x$cptable[min, "xerror"] 
}

hyper_grid %>%
  mutate(
    cp    = purrr::map_dbl(models, get_cp),
    error = purrr::map_dbl(models, get_min_error)
    ) %>%
  arrange(error) %>%
  top_n(-5, wt = error)



optimal_tree =  rpart(
    formula = sale_price ~ .,
    data    = train,
    method  = "anova",
    control = list(minsplit = 9, maxdepth = 13, cp = 0.01)
    )

rpart.plot(optimal_tree)
plotcp(optimal_tree)

optimal_tree$cptable

predictionrt <- predict(optimal_tree, newdata = test)
RMSE(predictionrt, test$sale_price)

#It seems even with the optimal model while the error is down by 1%-2% it is not no where near what we want I think I might need to rework our features, but lets try bagging

```


```{r}


set.seed(123)

# train bagged model
bagged_m1 <- bagging(
  formula = sale_price ~ .,
  data    = train,
  coob    = TRUE
)

ntree <- 10:50

# create empty vector to store OOB RMSE values
rmse <- vector(mode = "numeric", length = length(ntree))

for (i in seq_along(ntree)) {
  
  set.seed(123)
  
  # perform bagged model
  model <- bagging(
  formula = sale_price ~ .,
  data    = train,
  coob    = TRUE,
  nbagg   = ntree[i]
)
  # get OOB error
  rmse[i] = model$err
}

plot(ntree, rmse, type = 'l', lwd = 2)
abline(v = 25, col = "red", lty = "dashed")
ctrl = trainControl(method = "cv",  number = 10) 

# CV bagged model
cv_model = train(
  sale_price ~ .,
  data = train,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )

# assess results
cv_model


predictionbagrt = predict(cv_model, newdata = test)

RMSE(predictionbagrt, test$sale_price)

RTcomparePrediction = predict(cv_model, house_imp)
RMSE(RTcomparePrediction, house_imp$sale_price)

# plot most important variables
plot(varImp(cv_model), 20)  

#Well we reduced RMSE slightly to our original RMSE but still the results are poor I believe this was a better indicator of our model than the linear model where we had a better r2 

#I will be revising some of the features and re-try this with new set of features as this yielded poor results, I believe removing features with high missing data and low importance might yield to a better model like total_taxes, fuel_type, and etc
```

let's fit the linear model

```{r}
linear_mod = lm(sale_price ~ ., house_imp)
linear_mod$model
options(scipen = 999)
predictionlin = predict(linear_mod,house_imp)
predictionlin
RMSE(predictionlin, house_imp$sale_price)
#last_house_df = cbind(final_house_df[,1:15], final_house_df[, 17:19])
#glimpse(last_house_df)
summary(linear_mod)
sd(linear_mod$residuals)
summary(linear_mod)$sigma
summary(linear_mod)$r.squared
```

Lets try Random-Forest

```{r}
library(randomForest)
rf_model <- randomForest(sale_price~.,
                            data = train)

importance    <- importance(rf_model)
importance
varImpPlot(rf_model)

predictionrf <- predict(rf_model,test)
predictionrf
RMSE(predictionrf,test$sale_price)
#Well while they had similar RMSE it is still high while I would not ship this model to the real-world as you can see from the work clearly, either my data cleaning or feature selection was poor here, I have attempted and tweaked with multiple features but it seems I did not gain much significance in lowering RMSE


RFcomparePrediction = predict(rf_model, house_imp)
RMSE(RFcomparePrediction,house_imp$sale_price)
plot(house$sale_price, ylim = range(house$sale_price,predictionlin,predictionrt,predictionrf), col = 'red')
lines(predictionlin, col = 'yellow')
lines(predictionrt, col = 'brown')
lines(predictionrf, col = 'green')
df = data.frame(house_imp$sale_price, predictionlin, RTcomparePrediction, RFcomparePrediction)

require(scales)

ggplot(df, aes(x = df$predictionlin, y = df$house_imp.sale_price)) + geom_point() + labs(x = 'OLS Prediction', y = 'Actual Sale Price') + scale_x_continuous(labels = comma)
ggplot(df, aes(x = df$RTcomparePrediction, y = df$house_imp.sale_price)) + geom_point() + labs(x = 'Regression Tree Prediction', y = 'Actual Sale Price') + scale_x_continuous(labels = comma)
ggplot(df, aes(x = df$RFcomparePrediction, y = df$house_imp.sale_price)) + geom_point() + labs(x = 'RandomForest Prediction', y = 'Actual Sale Price') + scale_x_continuous(labels = comma)

# null_model_rmse = RMSE(sum(house$sale_price), mean(house$sale_price))
# null_model_rmse
# R2 <- 1 - (sum(((mean(house$sale_price))^2)/sum((house$sale_price-mean(house$sale_price))^2)))
# R2

sum(abs(house$sale_price - mean(house$sale_price))) / 528

null_model_r2
```